Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-29 WSSSPE1_report/WSSSPE1_summary_report.tex
Updating bibliography.
completed summaries in the frameworks area

\item Marlon Pierce, Suresh Marru, Chris Mattmann. Sustainable
  Cyberinfrastructure Software Through Open Governance~\cite{Pierce_WSSSPE}

@@

@@

>>>>

``paper considers how software sustainability can be addressed in the design and engineering process of software systems...it must be considered a composite, non- functional requirement, which can be reasoned over at an architectural level by building and evaluating appropriate architectural models.''

``Software Sustainability Institute define sustainability as ‘software you use today will be available - and continue to be improved and supported - in the future’ which implicitly suggests that sustainability is concerned with concepts of availability, extensibility, and the maintainability of the software [9]''

``software sustainability can be defined as ‘a measure of a systems extensibility, interoperability, maintainability, portability, reusability and scalability''

``For software to be sustainable it needs to be both useful and adaptable as stakeholders requirements, technology and environments evolve and change.''

``Architectural representations of systems are therefore effective in understanding broader system concerns by abstracting away from system details, hence the trend for reasoning about quality attributes at the architectural level''

They look to architectural level analysis to support sustainability, especially using scenario planning

They talk about sustainability going beyond the artifact itself, but then pivot to environmental sustainability.  Hard to see a framework coming from software engineering dealing with the range of contextual conditions relevant to whether and how a piece of software will continue to be improved and supported in the future.
They don't mention resource availability at all.
\item Marlon Pierce, Suresh Marru, Chris Mattmann. Sustainable
  Cyberinfrastructure Software Through Open Governance~\cite{Pierce_WSSSPE}

@@

Compares the data lifecycle with a software lifecycle, ``suggesting that the well­accepted and well­documented digital science data lifecycle management best practices may inform the development of sustainable scientific software.'' In particular they focus on adding metadata about software through its development, adding the idea of preservation (most likely via validation tests) to the software lifecycle, and assessing ``data in/data out'' connections.

@@

``providing some history and provocations for the use of an ecological metaphor to describe software development environments ... we want to ask if it can indeed be a productive and thought provoking one.''  They examine historical uses of the ecosystem metaphor (centered around feedback and self-regulation) across domains. They note the limits of the metaphor as it encounters human agency within the broad landscape of software work.  In essence they propose that study and best practices should (in addition to a focus on individual projects) focus on the relationships/dependencies/dynamics across and between projects/components.  In particular they argue for examining ``niches'' which sustain particular pieces of software, they define ``a software niche as the set of technical requirements, organizational conditions and cultural mores that support its maintenance and use over time.'' They call for better understanding and modeling of niches.
